Nested sampling with multiple scalars

### 2015-01-19 Paper/ms.tex
A bit of writing

Nested Sampling begins by drawing particles from the
prior $\pi(\xx)$ and successively imposing constraints on the value of
the likelihood $L(\xx)$ that compress the prior mass by a
factor that is approximately known.
This enables the calculation of the marginal likelihood
and properties of the posterior, or any other distribution that is intermediate
between the prior and the posterior.

@@

 For example, we might be interested in the
properties of the canonical distributions
$p(\mathbf{x} | \alpha, \beta) \propto \pi(\mathbf{x})\exp\left[\alpha f_1(\mathbf{x}) + \beta f_2(\mathbf{x})\right]$
for many different values of $\alpha$ and $\beta$, including the
normalisations or ``partition function'' $\mathcal{Z}(\alpha, \beta)$.
This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions.

If the prior is $\pi(\mathbf{x})$, then a family of canonical
probability distributions is given by
\begin{eqnarray}
p(\mathbf{x} | \lambda_1, ..., \lambda_n) = 
\frac{\pi(\mathbf{x})\exp(-\sum_{i=1}^n \lambda_i f_i(\mathbf{x}))}
{\mathcal{Z}(\lambda_1, ..., \lambda_n)}
\end{eqnarray}
This family of distributions arises from a MaxEnt update from the prior
$\pi(\mathbf{x})$ if the expected values of the functions
$\{f_i(\mathbf{x})\}$ are specified. In statistical mechanics it is useful to
quantify the properties of
$p(\mathbf{x} | \lambda_1, ..., \lambda_n)$ as a function of the
$\lambda$s.

\section{Nested sampling}

>>>>

Nested Sampling begins by drawing particles from the
prior $\pi(\xx)$ and successively imposing constraints on the value of
the likelihood $L(\xx)$ that compress the prior mass by a
factor that is approximately known.
This enables the calculation of the marginal likelihood
and properties of the posterior, or any other distribution that is intermediate
between the prior and the posterior. For example, we might be interested in
a ``power posterior'' where the likelihood is raised to a power $\beta$:
\begin{eqnarray}
p(\xx; \beta) &=& \frac{\pi(\xx)L(\xx)^\beta}{Z(\beta)}
\end{eqnarray}
The normalisation and posterior samples from this distribution can be obtained
from the original Nested Sampling run. An example application, in the case of
a Bayesian model with a ``gaussian noise'' assumption in the likelihood,
computing $p(\xx; \beta)$ for $\beta \neq 1$ allows us to explore what the
posterior distribution would have been if the noise variance had been greater.

@@

If we were only interested in a single canonical distribution, for example
with $\alpha_1 = 0.3$ and $\alpha_2 = 0.7$, we could estimate its normalising
constant and by running standard Nested Sampling with ``likelihood''
$L(\xx) = \exp\left[0.3S_1(\xx) + 0.7S_2(\xx)\right]$. However, usually we
are interested in a range of values for $\alpha_1$ and $\alpha_2$, and we
want to know the normalisation $Z(\alpha_1, \alpha_2)$, called the
partition function. This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions.

\section{Properties of Nested sampling}
We seek an algorithm for computing the partition function
$Z(\alpha_1, \alpha_2)$ for a range of values of the $\alpha$s, from a single
run. To be considered a variant of Nested Sampling, we require that the method
satisfies the following requirements:
\begin{itemize}
\item The algorithm should begin with $n$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$S_1(\xx)$ and $S_2(\xx)$ are greater than what would be expected from the
prior.
\item The algorithm should try to move through a sequence of probability
distributions whose enclosed prior masses are approximately known.
\item The algorithm should be invariant to monotonic transformations of
$S_1$ and $S_2$, i.e. it should only depend on rankings of $S_1$ and $S_2$
values and not the values themselves.
\end{itemize}
